Initial report

### 2022-12-07 preprocessing/initial_report.Rmd
created basic project structure and draft analysis for manuscript

---
title: "Initial report"
output: html_document
date: '`r Sys.Date()`'
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(viridis)
library(scales) # fto display percentages

@@

# Read data
raw <- readr::read_tsv("data/ECAW_raw_data.tsv")
# remove surveys where someone opened it, but didn't respond to any questions
raw <- raw %>% filter(progress!=0)
# number of respondents
nResponse <- raw %>% nrow()

# Filter for complete responses (I'm am maintaining the variable name "raw" only because it is used throughout the rest of the code)
raw <- raw %>% filter(finished)
#number of completed surveys
nResponseComplete <- raw %>% nrow()

invited <- 1148 # Number provided from Kyara at ALSPAC
bounced <- 54 # Number provided from Kyara at ALSPAC
received <- invited - bounced
```

# Response rate

We invited the ALSPAC mailing list to participate, which included `r invited` email addresses, of which `r bounced` bounced, leaving `r received` emails. The survey was completed `r nResponseComplete` times and partially completed `r nResponse - nResponseComplete` times. This gives a response rate for completed surveys of `r percent(nResponseComplete / received)` and a response rate for any questions `r percent(nResponse / received)`.

The median time taken for complete survey responses was `r (median(raw$duration_in_seconds) / 60) %>% round(1)` minutes (IQR: `r (quantile(raw$duration_in_seconds, .25) / 60) %>% round(1)` - `r (quantile(raw$duration_in_seconds, .75) / 60) %>% round(1)`)

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
typically_likert_data <-
  raw %>% 
    select(starts_with("typically_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("time", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
            "Strongly disagree",
            "Somewhat disagree",
            "Neither agree nor disagree",
            "Somewhat agree",
            "Strongly agree",
            "Missing",
            "I don't understand the question")
            ),
      type = str_to_title(type)
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

typically_missing_data <-
  typically_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

@@

# Current practices

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
method_data <-
  raw %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind the data analyst",
        type == "confirmatory" ~ "Contain confirmatory analysis",
        type == "exploratory" ~ "Contain exploratory analysis",
        type == "preregistered" ~ "Are preregistered",
        type == "script" ~ "Share analysis scripts",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

# ALSPAC and ECAWs

@@

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
raw %>% 
  select(n_studies) %>% 
  filter(!is.na(n_studies)) %>%
  ggplot() +
  aes(
    x = n_studies
  ) +
  geom_histogram() +
  labs(
    x = "The number of studies published",
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line()
  )
```

>>>>

---
title: "Initial report"
output: html_document
date: '`r Sys.Date()`'
editor_options: 
  chunk_output_type: console
---
TODO: subset by concerned number of studies and current methods
```{r setup, include=FALSE}
# General rmarkdown rendering setting
knitr::opts_chunk$set(echo = TRUE)

# Loading used packages 
library(tidyverse)
library(likert)
library(viridis)

@@

# Read data
raw <- readr::read_tsv("data/ECAW_raw_data.tsv")
```

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 20}
# Prepearing data for Likert-scale plotting
typically_likert_data <-
  raw %>% 
    select(starts_with("typically_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("time", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
            "Strongly disagree",
            "Somewhat disagree",
            "Neither agree nor disagree",
            "Somewhat agree",
            "Strongly agree",
            "Missing",
            "I don't understand the question")
            ),
      type = str_to_title(type)
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  ungroup()

# Creating text labels for reporting the count of non-likert response levels on the Likert-scale plot 
typically_missing_data <-
  typically_likert_data %>% 
  filter(scale %in% c("Missing", "I don't understand the question")) %>% 
  mutate(
    scale = case_when(
      scale == "I don't understand the question" ~ "Don't understand",
      scale == "Missing" ~ "Missing"),
    text = paste0(scale, ": ", n)
    ) %>% 
  group_by(type) %>% 
  summarise(text = stringr::str_c(text, collapse = "\n"))

@@

# ALSPAC

@@

# Current methods

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width = 10}
method_data <-
  raw %>% 
    select(starts_with("method_")) %>% 
    pivot_longer(
      cols = everything(),
      names_to = "variable",
      values_to = "scale",
    ) %>%
    separate(variable, into = c("name_prefix", "type")) %>% 
    group_by(type) %>% 
    count(scale) %>% 
  ungroup() %>% 
    replace_na(list(scale = "Missing")) %>% 
    mutate(
      scale = factor(scale, levels = c(
          "Never or almost never",
          "Sometimes",
          "About half the time",
          "Most of the time",
          "Always or almost always",
          "Missing",
          "I don't understand the question")
            ),
      type = case_when(
        type == "blind" ~ "Blind analysis",
        type == "confirmatory" ~ "Confirmatory analysis",
        type == "exploratory" ~ "Exploratory analysis",
        type == "preregistered" ~ "Preregistration",
        type == "script" ~ "Analysis script",
      )
      ) %>% 
  group_by(type) %>%
  tidyr::complete(scale, fill = list(n = 0)) %>% 
  mutate(
    percentage = n / sum(n)
  ) %>% 
  ungroup()

@@

```{r, echo = FALSE, message = FALSE, warning = FALSE}
# Separate zero and bins of 5
raw %>% 
  select(n_studies) %>% 
  filter(!is.na(n_studies)) %>%
  ggplot() +
  aes(
    x = n_studies
  ) +
  geom_histogram(binwidth = 3) +
  labs(
    x = "The number of studies published",
    y = "Count"
  ) +
  scale_y_continuous(expand = c(0, 0)) +
  scale_x_continuous(
    # breaks = seq(0, 100, by = 3)
    )
  theme(
    panel.background = element_blank(),
    panel.grid = element_blank(),
    axis.line = element_line()
  )
```
